---
title: "Which Bus Now?"
author: "Miles McBain"
date: "31 March 2017"
output: html_document
---
```{r}
library(readr)
library(tidyverse)
library(fields)
```


# Introduction
# Dependencies
# Reference Data
https://translink.com.au/about-translink/open-data
# Learning Outcomes

#Analysis

##Get the data. 
Assumed Present. Code to add from tslumley.

##Unzip data
```{r}
uzip <- utils::unzip("../static_data/SEQ_GTFS.zip")
```

##Import data we need
```{r}
stops <- read_csv("./stops.txt", guess_max = 10000) #a bit hacky
```

##Find a stop based on a query string
```{r}
#' Find a stop
#' @descrpion Find a stop based on a comma separated qury string of search terms:
#' eg: "Montague Rd, Kurilpa Street" to match "Montague Rd at Kurilpa Street"
#' @param stop_names_list a column of stop names
#' @param query the query string
#' @example stops %>% filter(find_stop(stop_name,"Montague Rd, Kurilpa Street"))

find_stop <- function(stop_names_list, query){
    query_pieces <- unlist(strsplit(query, split = "\\s|,"))
    query_pieces <- example[lapply(example, nchar) != 0] #drop the exmply matches
    regex_query <- case_when(example == "Rd" | example == "Road" ~ "(Rd|Road)",
                            example == "St" | example == "Street" ~ "(St|Street)",
                            example == "Ln" | example == "Lane" ~ "(Ln|Lane)",
                            TRUE ~ example) %>%
    paste0(collapse = ".*")
    grepl(x = stop_names_list, pattern = regex_query)#return the logical vector.
}

```

<<<<<<< HEAD
## Find the distance in km between two bus stops
```{r}
#' Calcuate a distance
#' @descrpion Calculate a distance as the crow flies between two bus stops
#' @param start bust stop as latitude and longitude 
#' @param stop bust stop as latitude and longitude 
#' @example distance_between_stops(1330, 1331)
distance_between_stops <- function(stop1, stop2){
  p1 <- stops %>%
    filter(stop_id == stop1) %>%
    select(stop_lat, stop_lon)
  p2 <- stops %>%
    filter(stop_id == stop2) %>%
    select(stop_lat, stop_lon)
  
    return(rdist.earth.vec(p1, p2, miles = FALSE))
  }
=======
## Find the next departing service
```{r}
stop_times <- read_csv("stop_times.txt", guess_max = , 
                       col_types = cols(
                          trip_id = col_character(),
                          arrival_time = col_character(),
                          departure_time = col_character(),
                          stop_id = col_character(),
                          stop_sequence = col_character(),
                          pickup_type = col_character(),
                          drop_off_type = col_character()
                       )
              )
#need to set 



#Has times greater than 24:00:00!
>>>>>>> 485f79d3291e03945fd340ae2911f6db192cb040
```

### Find the bus need to get to your destination
#' Find the bus need to get to your destination on a weekday
#' @descrpion Find the bus need to get to your destination on a weekday
#' @param dest end of journey as a stip id 
#' @param arrive_tm as an arival time
#' @return short name aka bus nbr and route_long_name
#' @example find_a_route_weekday(1880, "16:30:00")
```{r}
find_a_route_weekday <- function(dest, arrive_tm){
  ## Find a route
  # 1. stop_times.txt 
  # srch by bus id (1880 ) & time (16:30). Rtn Trip ID. E.G. 9092845-BT 16_17-FEB_FUL-Weekday-05
  bus <- stop_times %>%
    filter(stop_id == dest & departure_time == as.numeric(hms(arrive_tm)) & grepl('FUL-Weekday-05', trip_id)) %>%
    select(trip_id)
  
  # 2. trips.txt. 
  # srch by trip id (9092845-BT 16_17-FEB_FUL-Weekday-05). Rtn Route id. e.g. 66-738
  route <- trips %>%
    filter(trip_id == as.character(bus)) %>%
    select(route_id)
  
  #3. routes.txt
  #Srch by route id. Rtn route_short_name & route_long_name
  journey <- routes %>%
    filter(route_id == as.character(route)) %>%
    select(route_short_name, route_long_name)
  
  return(journey)
}
```

### Find all the stops along a route. Origin and Destination inclusive
#' Find all the stops along a route. Origin and Destination inclusive
#' @descrpion Find all the stops along a route. Origin and Destination inclusive
#' @param dest end of journey as a trip_id 
#' @return trip_id nbr and stop_id  as adata frame
#' @example find_journey_stops(bus)
```{r}

# Find all the stops along a route. Origin and Destination inclusive
find_journey_stops <- function(trip){
  journey_stops <- stop_times %>%
    filter(trip_id == as.character(trip)) %>%
    select(trip_id, stop_id)
  
  return(journey_stops)
}
```

