---
title: "Interactive bus maps"
author: "Thomas Lumley"
date: "31 March 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

An interactive map doesn't fit as well into an Rmd notebook: we'll create an HTML widget instead
First, we unzip the supplied file into a temporary directory. The temporary direction will be deleted when you quit R.
```{r}
staticGTFS<-tempdir()
unzip("../static_data/SEQ_GTFS.zip",exdir=staticGTFS) #FIXME will use package.file() eventually
list.files(staticGTFS)
```

```{r}
stops<-read.csv(paste(staticGTFS,"stops.txt",sep="/"), stringsAsFactors = FALSE)
summary(stops)
stops$inzone<-as.integer(substr(stops$zone_id,1,1))
```

We'll also get a count of buses for each stop can plot the stops to get the crudest possible map. This is fairly slow because there are millions of trip/stop combinations in `stop_times.txt`.  We'll also filter to just the services on 31 March 2017, which was a Friday.
```{r}
stop_times <- read.csv(paste(staticGTFS,"stop_times.txt",sep="/"), stringsAsFactors = FALSE)
today<-20170331
calendar <- read.csv(paste(staticGTFS,"calendar.txt",sep="/"), stringsAsFactors = FALSE)
calendar <-subset(calendar, start_date <= today & end_date>=today)
exceptions<- read.csv(paste(staticGTFS,"calendar_dates.txt",sep="/"), stringsAsFactors = FALSE)
if (!any(exceptions$date==today)) cat("No exceptions today, fortunately\n")
services_today<-calendar$service_id[calendar$friday==1]
trips <- read.csv(paste(staticGTFS,"trips.txt",sep="/"))
trips<- trips[trips$service_id %in% services_today,]
trip_route <- merge(stop_times, trips, by="trip_id")
bus_count<-with(trip_route,by(trip_id,list(stop=stop_id),length))
head(bus_count)
bus_count_df<-data.frame(stop_id=names(bus_count), counts=as.vector(bus_count))
fri_stops<-merge(stops, bus_count_df,by="stop_id")
```

Now, there's documentation for the `leaflet` package at [RStudio](https://rstudio.github.io/leaflet/), and we're going to use `htmlwidgets` to save the interactive map in a file.

```{r}
library(leaflet)
library(htmlwidgets)
fri_stops$label<-with(fri_stops, paste(stop_name,counts,paste("stop",stop_id),sep="<br>"))
busmap <- leaflet(fri_stops) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(lng=~stop_lon,lat=~stop_lat, radius=~I(round(2+sqrt(counts/10))),
                   col=~I(palette()[inzone]),stroke=FALSE, popup=~label)
saveWidget(busmap, file='/tmp/buses/buses.html',selfcontained=FALSE)
```

And, for comparison, Saturday April 1

```{r}
stop_times <- read.csv(paste(staticGTFS,"stop_times.txt",sep="/"), stringsAsFactors = FALSE)
tomorrow<-20170401
calendar <- read.csv(paste(staticGTFS,"calendar.txt",sep="/"), stringsAsFactors = FALSE)
calendar <-subset(calendar, start_date <= tomorrow & end_date>=tomorrow)
exceptions<- read.csv(paste(staticGTFS,"calendar_dates.txt",sep="/"), stringsAsFactors = FALSE)
if (!any(exceptions$date==tomorrow)) cat("No exceptions today, fortunately\n")
services_today<-calendar$service_id[calendar$saturday==1]
trips <- read.csv(paste(staticGTFS,"trips.txt",sep="/"))
trips<- trips[trips$service_id %in% services_today,]
trip_route <- merge(stop_times, trips, by="trip_id")
bus_count<-with(trip_route,by(trip_id,list(stop=stop_id),length))
head(bus_count)
bus_count_df<-data.frame(stop_id=names(bus_count), counts=as.vector(bus_count))
sat_stops<-merge(stops, bus_count_df,by="stop_id")
```

And render in leaflet:

```{r}
library(leaflet)
library(htmlwidgets)
sat_stops$label<-with(sat_stops, paste(stop_name,counts,paste("stop",stop_id),sep="<br>"))
busmap <- leaflet(sat_stops) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(lng=~stop_lon,lat=~stop_lat, radius=~I(round(2+sqrt(counts/10))),
                   col=~I(palette()[inzone]),stroke=FALSE, popup=~label)
saveWidget(busmap, file='/tmp/buses/buses_saturday.html',selfcontained=FALSE)
```

And now a direct map of the differences in counts.
```{r}
diff_stops<-merge(sat_stops,fri_stops,by="stop_id")
diff_stops$diff<-with(diff_stops, counts.y-counts.x)
diff_stops$dlabel<-with(diff_stops, paste(stop_name.x,paste("Sat:",counts.x),paste("Fri:",counts.y),sep="<br>"))
busmap <- leaflet(diff_stops) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(lng=~stop_lon.x,lat=~stop_lat.x, radius=~I(sqrt(abs(diff))),
                   col=~I(ifelse(diff< 0,"blue","orange")),stroke=FALSE, popup=~dlabel)
saveWidget(busmap, file='/tmp/buses/buses_diff.html',selfcontained=FALSE)
```

As we'd expect, the map is mostly orange, with very few stops having more buses on Saturday than Friday. However, a few do, such as "Montague Rd at Davies Park, stop 8", which has a Saturday market. 